### 2013-07-24 doc/asplos14/method.tex
fix punctuation

\vspace{0.5ex}
\noindent{\bf Unbalanced tree search in-memory (UTS-Mem):} Unbalanced Tree Search
(UTS) is a benchmark for evaluating the programmability and performance of
systems for parallel applications that require dynamic load
balancing~\cite{UTS}. It involves traversing an unbalanced implicit tree: at
each vertex, its number of children is sampled from some probability
distribution, and this number of new nodes are added to a work queue to be
visited. While this benchmark captures irregular, dynamic \emph{computation,}
we actually want to evaluate performance of algorithms with irregular
\emph{memory\/} access patterns. Thus we augment UTS by using the existing
traversal code to create a large tree in memory, and then we traverse the
in-memory tree. In our modified UTS, the timed portion is this
traversal of the in-memory tree. This in-memory traversal has no knowledge of
the tree structure beforehand.

\vspace{0.5ex}
\noindent{\bf Breadth-first-search (BFS):} This is the primary kernel for the
Graph500 benchmark and is what currently determines the ranking of machines on
the Graph500 list~\cite{graph500list}. As a whole, the Graph500 benchmark
suite is designed to bring the focus of system design on data-intensive
workloads, particularly large-scale graph analysis problems, that are
important among cybersecurity, informatics, and network-understanding
workloads. The BFS benchmark builds a search tree containing parent nodes for
each traversed vertex during the search. While this is a relatively simple
problem to solve, it exercises the random-access and fine-grained
synchronization capabilities of a system as well as being a primitive in many
other graph algorithms. Performance is measured in \emph{traversed edges per
second\/} (TEPS), where the number of edges is the edges making up the
generated BFS tree. With some modifications to the XMT reference version of Graph500 BFS,
the XMT compiler can be made to recognize and apply a Manhattan loop collapse, exposing enough parallelism to allow it to scale out to 64 nodes for the problem scales we show.
In order to make comparison easier, we do not employ algorithmic improvements for any of these versions, though there are
many~\cite{Beamer:Graph500,Yoo:FixedPointGraph500}; this makes our results difficult to compare with published Graph500 results. \Grappa can be expected to benefit the same as MPI due to decreased communication.

\vspace{0.5ex}
\noindent{\bf IntSort:} This sorting benchmark is taken from the NAS Parallel Benchmark Suite~\cite{Bailey91thenas,nas3.3} and is one on which the Cray XMT's early predecessor once held the world speed record~\cite{TeraRecord}. The largest problem size, class D, ranks two billion uniformly distributed random integers using either a bucket or a counting sort algorithm, depending on the strengths of the system. Bucket sort executes a greater number of loops, but is able to leverage locality and avoid communication completely in the final phase, ranking within buckets. For these reasons, the MPI reference version and our \Grappa implementation use bucket sort. On the other hand, the Cray XMT cannot take advantage of locality, but has an efficient compiler-supported parallel prefix sum, so it performs best using the counting sort algorithm.  The performance metrics for NAS Parallel Benchmarks, including IntSort, are ``millions of operations per second'' (MOPS). For IntSort, this ``operation'' is ranking a single key, so it is roughly comparable to ``GUPS'' or ``TEPS.''

\vspace{0.5ex}
\noindent{\bf PageRank:} This is a common centrality metric for
graphs. PageRank is an iterative algorithm with a common
pattern of gather, apply, and scatter on the rank of vertex. The algorithm
is often implemented by sparse linear algebra libraries, with the main
kernel being the sparse matrix dense vector multiply. For the multiply
step, \Grappa parallelizes over the rows and parallelizes each dot
product. PageRank has the fortunate property that the accumulation
function over the in-edges is associative and commutative, so they can
be processed in any order or in parallel. Rather than the programmer writing the
parallel dot product as local accumulations with a final all-reduce
step, we simply send streaming increments to each element of the final
vector. We compare PageRank to published results for the Trilinos linear algebra library implemented in MPI~\cite{Plimpton:2011:MML:2286659.2286704},
 and multithreaded PageRank for the XMT~\cite{MTGL:pagerank}.
For \Grappa, we run on a scale 29 graph using the Graph500 generator.

>>>>

\vspace{0.5ex}
\noindent{\bf Unbalanced tree search in-memory (UTS-Mem)} Unbalanced Tree Search
(UTS) is a benchmark for evaluating the programmability and performance of
systems for parallel applications that require dynamic load
balancing~\cite{UTS}. It involves traversing an unbalanced implicit tree: at
each vertex, its number of children is sampled from some probability
distribution, and this number of new nodes are added to a work queue to be
visited. While this benchmark captures irregular, dynamic \emph{computation,}
we actually want to evaluate performance of algorithms with irregular
\emph{memory\/} access patterns. Thus we augment UTS by using the existing
traversal code to create a large tree in memory, and then we traverse the
in-memory tree. In our modified UTS, the timed portion is this
traversal of the in-memory tree. This in-memory traversal has no knowledge of
the tree structure beforehand.

\vspace{0.5ex}
\noindent{\bf Breadth-first-search (BFS)} This is the primary kernel for the
Graph500 benchmark and is what currently determines the ranking of machines on
the Graph500 list~\cite{graph500list}. As a whole, the Graph500 benchmark
suite is designed to bring the focus of system design on data-intensive
workloads, particularly large-scale graph analysis problems, that are
important among cybersecurity, informatics, and network-understanding
workloads. The BFS benchmark builds a search tree containing parent nodes for
each traversed vertex during the search. While this is a relatively simple
problem to solve, it exercises the random-access and fine-grained
synchronization capabilities of a system as well as being a primitive in many
other graph algorithms. Performance is measured in \emph{traversed edges per
second\/} (TEPS), where the number of edges is the edges making up the
generated BFS tree. With some modifications to the XMT reference version of Graph500 BFS,
the XMT compiler can be made to recognize and apply a Manhattan loop collapse, exposing enough parallelism to allow it to scale out to 64 nodes for the problem scales we show.
In order to make comparison easier, we do not employ algorithmic improvements for any of these versions, though there are
many~\cite{Beamer:Graph500,Yoo:FixedPointGraph500}; this makes our results difficult to compare with published Graph500 results. \Grappa can be expected to benefit the same as MPI due to decreased communication.

\vspace{0.5ex}
\noindent{\bf IntSort} This sorting benchmark is taken from the NAS Parallel Benchmark Suite~\cite{Bailey91thenas,nas3.3} and is one on which the Cray XMT's early predecessor once held the world speed record~\cite{TeraRecord}. The largest problem size, class D, ranks two billion uniformly distributed random integers using either a bucket or a counting sort algorithm, depending on the strengths of the system. Bucket sort executes a greater number of loops, but is able to leverage locality and avoid communication completely in the final phase, ranking within buckets. For these reasons, the MPI reference version and our \Grappa implementation use bucket sort. On the other hand, the Cray XMT cannot take advantage of locality, but has an efficient compiler-supported parallel prefix sum, so it performs best using the counting sort algorithm.  The performance metrics for NAS Parallel Benchmarks, including IntSort, are ``millions of operations per second'' (MOPS). For IntSort, this ``operation'' is ranking a single key, so it is roughly comparable to ``GUPS'' or ``TEPS.''

\vspace{0.5ex}
\noindent{\bf PageRank} This is a common centrality metric for
graphs. PageRank is an iterative algorithm with a common
pattern of gather, apply, and scatter on the rank of vertex. The algorithm
is often implemented by sparse linear algebra libraries, with the main
kernel being the sparse matrix dense vector multiply. For the multiply
step, \Grappa parallelizes over the rows and parallelizes each dot
product. PageRank has the fortunate property that the accumulation
function over the in-edges is associative and commutative, so they can
be processed in any order or in parallel. Rather than the programmer writing the
parallel dot product as local accumulations with a final all-reduce
step, we simply send streaming increments to each element of the final
vector. We compare PageRank to published results for the Trilinos linear algebra library implemented in MPI~\cite{Plimpton:2011:MML:2286659.2286704},
 and multithreaded PageRank for the XMT~\cite{MTGL:pagerank}.
For \Grappa, we run on a scale 29 graph using the Graph500 generator.
